---
title: "Carnac Menhirs Stadium Profitability Decision Support"
subtitle: "Classification Modelling and Operational Scenario Analysis"
author: "Md Soad Solaiman"
date: "15 July 2025"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: false
    theme: readable
    css: styles.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  fig.width = 8.5,
  fig.height = 5.25,
  dpi = 120
)
```

```{r packages, include=FALSE}
library(tidyverse)
library(tidymodels)
library(rpart.plot)
library(vip)
library(car)
library(yardstick)
library(kableExtra)
library(scales)
```

```{r scroll-table-helper, include=FALSE}
# Wraps a kable table in a fixed-height scroll box, so long or wide
# tables stay contained instead of stretching the page.
scroll_table <- function(kbl, height = "220px") {
  kbl %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                  full_width = FALSE) %>%
    scroll_box(width = "100%", height = height)
}

plain_table <- function(kbl) {
  kbl %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                  full_width = FALSE)
}
```

# 1. Executive Summary

Carnac Menhirs has purchased an old basketball stadium and renamed it La Stade des Menhirs. The managing committee needs to know whether the current first-season configuration is likely to be profitable, and whether one practical operational change could improve that outlook.

The analysis fits a classification tree and a logistic regression model to 25,000 comparable club-season records. The current scenario, 40 combined years of experience across seven players, 1,500 members, poor stadium condition, no recent premiership and a kiosk, is classified as **not profitable** by both models. The logistic model estimates a 41.78% probability of profit.

Changing the café to a vending machine lowers that estimate to 31.66%. Changing it to a bar raises it to 54.41%, and both models classify the bar scenario as profitable. This is a modelled scenario comparison, not a guarantee of financial return, since conversion and operating costs are not included in either model.

# 2. Business Context

Carnac Menhirs has assembled a new seven-player team with 40 combined years of experience. The club's home venue, La Stade des Menhirs, is in poor condition. 1,500 members have joined for the coming season. The proposed spectator-service format is a kiosk.

The committee wants two things: evidence on first-season profitability risk, and a single operational change that could improve the modelled outcome. The intended audience is non-technical, so results below are interpreted in operational and financial terms throughout.

# 3. Decision Problem

This analysis addresses three management decisions:

- Is the current stadium configuration predicted to be profitable?
- Which club and venue factors matter most to that prediction?
- Does changing the café format improve the probability enough to justify a detailed commercial assessment?

The models classify profitability and compare scenarios. They do not estimate the size of a profit or loss, and they do not account for conversion costs.

# 4. Business Questions

- How should profit be defined for comparable stadium-owning clubs?
- Which operational variables matter most in a classification tree?
- Which individual factors and interactions matter in the logistic model?
- How accurately does the selected model classify held-out records?
- What do sensitivity, specificity, ROC and AUC say about model reliability?
- Is La Stade des Menhirs predicted to make a first-season profit?
- How do vending-machine and bar scenarios change the predicted outcome?
- Which tested intervention offers the strongest modelled improvement?

# 5. Dataset

The dataset contains 25,000 club-season records and seven source variables: revenue and costs (in thousands of dollars), average years of player experience, paying membership, premiership history over the previous 10 years, stadium condition, and café type.

```{r load-data, include=FALSE}
stades <- read_csv("../data/stades_2.csv", show_col_types = FALSE)
n_raw <- nrow(stades)
```

```{r show-raw-sample}
knitr::kable(
  head(stades, 3),
  caption = "Source club-season records"
) %>%
  scroll_table(height = "180px")
```

# 6. Data Preparation

The source data is pre-cleaned. All `r n_raw` records are preserved. A fixed seed keeps the row order reproducible, and variable names and types are standardised.

```{r reproducible-ordering, include=FALSE}
set.seed(1978974)
stades <- sample_n(stades, 25000, replace = FALSE)
```

```{r standardise-variables, include=FALSE}
stades <- stades %>%
  rename(
    revenue = Revenue,
    costs = Costs,
    ade = `a d'e`,
    members = MEMBERS,
    coupe = COUPE,
    cond = Cond,
    cafe = Cafe
  ) %>%
  mutate(
    members = as.numeric(members),
    ade = as.numeric(ade),
    revenue = as.numeric(revenue),
    costs = as.numeric(costs),
    coupe = as.factor(coupe),
    cond = as.factor(cond),
    cafe = as.factor(cafe)
  )
```

```{r show-standardised-sample}
knitr::kable(
  head(stades, 5),
  caption = "Standardised club-season records (sample)"
) %>%
  scroll_table(height = "200px")
```

# 7. Profit Definition

Profit is classified as `yes` when revenue exceeds costs, and `no` otherwise. Revenue and costs are then removed from the predictor set, since they directly determine the response and would leak the outcome into the models.

```{r define-profit, include=FALSE}
stades <- stades %>%
  mutate(profit = ifelse(revenue > costs, "yes", "no")) %>%
  mutate(profit = factor(profit, levels = c("no", "yes"))) %>%
  select(profit, everything()) %>%
  select(-revenue, -costs)

profit_counts <- count(stades, profit)
n_profit_yes <- profit_counts$n[profit_counts$profit == "yes"]
n_profit_no  <- profit_counts$n[profit_counts$profit == "no"]
```

Of the `r nrow(stades)` records, `r n_profit_yes` are classified as profitable and `r n_profit_no` as not profitable.

```{r show-profit-sample}
knitr::kable(
  head(stades, 5),
  caption = "Modelling data after profit classification (sample)"
) %>%
  scroll_table(height = "200px")
```

# 8. Classification Tree

The classification tree uses every remaining variable to predict profit.

```{r classification-tree, include=FALSE}
stades_tree <- decision_tree(mode = "classification") %>%
  set_engine("rpart") %>%
  fit(profit ~ ., data = stades)
```

```{r show-tree-plot}
rpart.plot(
  stades_tree$fit,
  extra = 1,
  type = 2,
  roundint = FALSE,
  main = "Classification Tree for Stadium Profitability"
)
```

The tree contains four splits and five terminal nodes. Each terminal node represents a segment of club-season records assigned a predicted profit class.

```{r variable-importance}
stades_tree %>%
  vip(num_features = 5) +
  aes(fill = Variable) +
  scale_fill_manual(values = c("#1B2A4A", "#C8963E", "#2C5282", "#B23B3B", "#2F7D5A")) +
  theme_minimal() +
  theme(legend.position = "none") +
  labs(
    title = "Variable Importance for Stadium Profitability",
    subtitle = "Classification-tree predictor ranking",
    x = "Importance",
    y = "Predictor"
  )
```

Premiership history, membership and café type are the most important predictors in the tree's variable-importance ranking.

# 9. Logistic Regression

## 9.1 Training and testing data

```{r train-test-split, include=FALSE}
set.seed(1978974)
pro_por <- 18000 / 25000

stades_split <- initial_split(stades, prop = pro_por)
stades_train <- training(stades_split)
stades_test <- testing(stades_split)
```

The reproducible split assigns `r nrow(stades_train)` records to training and `r nrow(stades_test)` to testing.

## 9.2 Baseline logistic regression

```{r baseline-logistic-model, include=FALSE}
stades_fit <- logistic_reg() %>%
  set_engine("glm") %>%
  fit(profit ~ ., data = stades_train)
```

The baseline model includes average experience, membership, premiership history, stadium condition and café type. Stadium-condition terms do not meet the 93% significance threshold used throughout this analysis.

# 10. Interaction Analysis

The interaction model includes all individual predictors and all second-order interactions between them.

```{r interaction-model, include=FALSE}
stades_lin_fit <- logistic_reg() %>%
  set_engine("glm") %>%
  fit(profit ~ .^2, data = stades_train)

interaction_anova_df <- Anova(stades_lin_fit$fit) %>%
  as.data.frame() %>%
  rownames_to_column("Term") %>%
  mutate(across(where(is.numeric), ~ signif(.x, 4)))
```

```{r show-interaction-anova}
knitr::kable(
  interaction_anova_df,
  caption = "Type II analysis of deviance, interaction model"
) %>%
  plain_table()
```

`members:coupe` and `members:cafe` are the only interactions meeting the 93% significance threshold, with p-values of 4.536e-06 and 3.266e-08 respectively.

# 11. Model Selection

Interactions with average experience are excluded, while experience is retained as an individual predictor. Backward elimination then proceeds at a 93% significance level, a stricter removal threshold (0.07) than the conventional 95% level.

```{r reduced-interaction-model, include=FALSE}
stades_rmv_exp <- logistic_reg() %>%
  set_engine("glm") %>%
  fit(
    profit ~ ade + members + coupe + cond + cafe +
      members:coupe + members:cond + members:cafe +
      coupe:cond + coupe:cafe + cond:cafe,
    data = stades_train
  )
```

`coupe:cafe` is removed first, with a p-value of 0.6608, the largest term above the 0.07 threshold.

```{r first-backward-step, include=FALSE}
stades_bsr <- logistic_reg() %>%
  set_engine("glm") %>%
  fit(
    profit ~ ade + members + coupe + cond + cafe +
      members:coupe + members:cond + members:cafe +
      coupe:cond + cond:cafe,
    data = stades_train
  )
```

Stadium condition and its remaining interactions are removed next, since none meet the 93% threshold. The final model retains average experience, membership, premiership history, café type, `members:coupe` and `members:cafe`. Main effects belonging to retained interactions stay in the model, preserving model hierarchy.

```{r final-logistic-model, include=FALSE}
stades_bsr <- logistic_reg() %>%
  set_engine("glm") %>%
  fit(
    profit ~ ade + members + coupe + cafe +
      members:coupe + members:cafe,
    data = stades_train
  )

final_anova_df <- as.data.frame(anova(stades_bsr$fit)) %>%
  rownames_to_column("Term") %>%
  mutate(across(where(is.numeric), ~ signif(.x, 4)))

final_coef_df <- tidy(stades_bsr$fit) %>%
  mutate(across(where(is.numeric), ~ signif(.x, 4)))
```

```{r show-final-anova}
knitr::kable(
  final_anova_df,
  caption = "Sequential analysis of deviance, final model"
) %>%
  plain_table()
```

```{r show-final-coefficients}
knitr::kable(
  final_coef_df,
  caption = "Final model coefficients"
) %>%
  plain_table()
```

In the final model, `members:coupe` and `members:cafe` have p-values of 1.411e-05 and 1.557e-08 respectively. Backward selection is sample-sensitive, and statistical significance does not necessarily imply operational importance. This specification is the completed analytical workflow, not a uniquely correct model.

## 11.1 Final logistic equation

Let \(\hat p_i\) denote the estimated probability that observation \(i\) is profitable. The fitted equation is:

$$
\log\!\left(\frac{\hat p_i}{1-\hat p_i}\right)
= -1.33
+ 0.136 \cdot ade_i
+ 0.000489 \cdot members_i
+ 0.860 \cdot \{\text{coupe}_i=\text{yes}\}
- 0.318 \cdot \{\text{cafe}_i=\text{kiosk}\}
- 0.685 \cdot \{\text{cafe}_i=\text{machine}\}
+ 0.000135 \cdot members_i \cdot \{\text{coupe}_i=\text{yes}\}
- 0.000127 \cdot members_i \cdot \{\text{cafe}_i=\text{kiosk}\}
- 0.000174 \cdot members_i \cdot \{\text{cafe}_i=\text{machine}\}
$$

The baseline categories are `coupe = "no"` and `cafe = "bar"`.

# 12. Held-Out Evaluation

## 12.1 Held-out predictions

```{r held-out-predictions, include=FALSE}
preds <- predict(stades_bsr, new_data = stades_test, type = "class")
stades_preds_testing <- predict(
  stades_bsr,
  new_data = stades_test,
  type = "prob"
)

classification_threshold <- 0.50

threshold_class <- factor(
  if_else(
    stades_preds_testing$.pred_yes >= classification_threshold,
    "yes",
    "no"
  ),
  levels = levels(stades_test$profit)
)

stopifnot(all(threshold_class == preds$.pred_class))

stades_evaluation <- bind_cols(
  stades_test %>% select(ade, members, coupe, cond, cafe, profit),
  preds %>% rename(pred_profit = .pred_class),
  stades_preds_testing
)
```

```{r show-held-out-sample}
knitr::kable(
  head(
    stades_evaluation %>%
      select(profit, pred_profit, .pred_no, .pred_yes),
    5
  ),
  caption = "Held-out profit classifications and probabilities (sample)"
) %>%
  scroll_table(height = "200px")
```

# 13. Confusion Matrix

The truth variable is `profit`. The predicted class is `pred_profit`. `.pred_yes` is the predicted probability of the profitable class. Both truth and predicted class use factor levels `no`, then `yes`, so `event_level = "second"` explicitly sets `yes` as the business event. The 0.50 threshold used for the class prediction is verified above.

```{r confusion-matrix, include=FALSE}
confusion_mat <- conf_mat(
  stades_evaluation,
  truth = profit,
  estimate = pred_profit
)

confusion_counts <- tibble(
  outcome = c(
    "True positive",
    "True negative",
    "False positive",
    "False negative"
  ),
  count = c(
    confusion_mat$table["yes", "yes"],
    confusion_mat$table["no", "no"],
    confusion_mat$table["yes", "no"],
    confusion_mat$table["no", "yes"]
  )
)

tp <- confusion_counts$count[confusion_counts$outcome == "True positive"]
tn <- confusion_counts$count[confusion_counts$outcome == "True negative"]
fp <- confusion_counts$count[confusion_counts$outcome == "False positive"]
fn <- confusion_counts$count[confusion_counts$outcome == "False negative"]
```

```{r show-confusion-counts}
knitr::kable(
  confusion_counts,
  caption = "Confusion-matrix counts, profitable records as the event"
) %>%
  plain_table()
```

`yardstick::conf_mat()` places predicted classes in rows and truth classes in columns. The held-out results are `r comma(tp)` true positives, `r comma(tn)` true negatives, `r comma(fp)` false positives and `r comma(fn)` false negatives.

```{r classification-metrics, include=FALSE}
classification_metrics <- bind_rows(
  yardstick::accuracy(stades_evaluation, truth = profit, estimate = pred_profit),
  yardstick::sens(stades_evaluation, truth = profit, estimate = pred_profit, event_level = "second"),
  yardstick::spec(stades_evaluation, truth = profit, estimate = pred_profit, event_level = "second"),
  yardstick::precision(stades_evaluation, truth = profit, estimate = pred_profit, event_level = "second"),
  yardstick::bal_accuracy(stades_evaluation, truth = profit, estimate = pred_profit, event_level = "second")
)

get_metric <- function(name) {
  round(classification_metrics$.estimate[classification_metrics$.metric == name] * 100, 2)
}

acc_pct  <- get_metric("accuracy")
sens_pct <- get_metric("sens")
spec_pct <- get_metric("spec")
prec_pct <- get_metric("precision")
bal_pct  <- get_metric("bal_accuracy")
```

```{r show-classification-metrics}
knitr::kable(
  classification_metrics,
  digits = 4,
  caption = "Held-out classification metrics at a 0.50 threshold"
) %>%
  plain_table()
```

Held-out accuracy is `r acc_pct`%. Sensitivity, or recall for profitable records, is `r sens_pct`%. Specificity for not-profitable records is `r spec_pct`%. Precision for profitable predictions is `r prec_pct`%. Balanced accuracy is `r bal_pct`%.

An earlier draft of this analysis labelled 21.56% as the sensitivity value. That figure is \(896/(3259+896)\), which is the false-negative rate, not sensitivity, once profitable records are set as the event. The underlying predictions are unchanged. Only the metric definition and label are corrected here.

## 13.1 Classification Threshold

The model converts probabilities to classes using the conventional 0.50 threshold. This default is not automatically optimal for the business decision. Lowering it would generally catch more potentially profitable stadiums, at the cost of more false positives. Raising it would reduce false positives, at the cost of missing more profitable stadiums. A different threshold should be chosen using business costs and validation data, not optimised on this held-out test set and then reported as unbiased performance.

# 14. ROC and AUC

```{r authoritative-roc, include=FALSE}
profit_roc <- yardstick::roc_curve(
  stades_evaluation,
  truth = profit,
  .pred_yes,
  event_level = "second"
)

profit_auc <- yardstick::roc_auc(
  stades_evaluation,
  truth = profit,
  .pred_yes,
  event_level = "second"
)

auc_pct <- round(profit_auc$.estimate * 100, 2)
```

```{r show-roc-plot}
autoplot(profit_roc) +
  geom_ribbon(
    aes(x = 1 - specificity, ymin = 0, ymax = sensitivity),
    fill = "#2C5282",
    alpha = 0.15
  ) +
  geom_path(
    aes(x = 1 - specificity, y = sensitivity),
    colour = "#1B2A4A",
    linewidth = 1.1
  ) +
  geom_abline(
    intercept = 0,
    slope = 1,
    linetype = "dashed",
    colour = "#B23B3B",
    linewidth = 0.6
  ) +
  labs(
    title = "ROC Curve for Held-Out Profit Classification",
    subtitle = "Event: profitable (`yes`); probability: `.pred_yes`",
    x = "1 - specificity",
    y = "Sensitivity"
  ) +
  theme_minimal()
```

With `profit` as truth, `yes` as the event and `.pred_yes` as the profitable-class probability, the AUC is `r auc_pct`%. It means the model has a 75.28 percent likelihood of correctly interpreting a randomly picked profitable record over an unprofitable one.

# 15. Stadium Scenarios

## 15.1 Current configuration

The current scenario of the club is that it has seven player. Their combined experience is 40 years. Their stadium condition is poor. It has 1500 members. And it is running a kiosk in their stadium.

```{r current-stadium-scenario}
new_tibble_team <- tibble(
  ade = 40 / 7,
  cond = "poor",
  coupe = "no",
  cafe = "kiosk",
  members = 1500
)

test_client <- predict(stades_bsr, new_data = new_tibble_team, type = "class")
pred_test_client <- predict(stades_bsr, new_data = new_tibble_team, type = "prob")

current_class <- as.character(test_client$.pred_class)
current_prob  <- round(pred_test_client$.pred_yes * 100, 2)



stades_tree_1 <- decision_tree(mode = "classification") %>%
  set_engine("rpart") %>%
  fit(profit ~ ., data = stades_train)

new_tree <- predict(stades_tree_1, new_data = new_tibble_team, type = "class")
new_tree_prob <- predict(stades_tree_1, new_data = new_tibble_team, type = "prob")

current_tree_class <- as.character(new_tree$.pred_class)
current_tree_prob  <- round(new_tree_prob$.pred_yes * 100, 2)



tibble(
  Model = c("Logistic regression", "Classification tree"),
  Prediction = c(current_class, current_tree_class),
  `Probability of profit` = c(paste0(current_prob, "%"), paste0(current_tree_prob, "%"))
) %>%
  knitr::kable(caption = "Current scenario prediction") %>%
  plain_table()

```

The logistic model classifies the current scenario as **`r current_class`**, with a `r current_prob`% probability of profit.

```{r scenario-classification-tree, include=FALSE}
stades_tree_1 <- decision_tree(mode = "classification") %>%
  set_engine("rpart") %>%
  fit(profit ~ ., data = stades_train)

new_tree <- predict(stades_tree_1, new_data = new_tibble_team, type = "class")
current_tree_class <- as.character(new_tree$.pred_class)
current_tree_class
```

```{r show-training-tree-plot}
rpart.plot(
  stades_tree_1$fit,
  extra = 1,
  type = 2,
  roundint = FALSE,
  main = "Training-Data Classification Tree"
)
```

The classification tree also predicts **`r current_tree_class`** for the current scenario.

## 15.2 Vending-machine scenario

```{r vending-machine-scenario}
new_tibble_team_3 <- tibble(
  ade = 40 / 7,
  cond = "poor",
  coupe = "no",
  cafe = "machine",
  members = 1500
)

pred_test_client_3 <- predict(stades_bsr, new_data = new_tibble_team_3, type = "prob")
class_test_client_3 <- predict(stades_bsr, new_data = new_tibble_team_3, type = "class")
new_tree_2 <- predict(stades_tree_1, new_data = new_tibble_team_3, type = "class")
new_tree_2_prob <- predict(stades_tree_1, new_data = new_tibble_team_3, type = "prob")

vending_prob  <- round(pred_test_client_3$.pred_yes * 100, 2)
vending_class <- as.character(class_test_client_3$.pred_class)
vending_tree_class <- as.character(new_tree_2$.pred_class)
vending_tree_prob  <- round(new_tree_2_prob$.pred_yes * 100, 2)

tibble(
  Model = c("Logistic regression", "Classification tree"),
  Prediction = c(vending_class, vending_tree_class),
  `Probability of profit` = c(paste0(vending_prob, "%"), paste0(vending_tree_prob, "%"))
) %>%
  knitr::kable(caption = "Vending-machine scenario prediction") %>%
  plain_table()
```

Switching the café to a vending machine lowers the estimated probability of profit to `r vending_prob`%.

## 15.3 Bar scenario

```{r bar-scenario}
new_tibble_team_2 <- tibble(
  ade = 40 / 7,
  cond = "poor",
  coupe = "no",
  cafe = "bar",
  members = 1500
)

pred_test_client_2 <- predict(stades_bsr, new_data = new_tibble_team_2, type = "prob")
class_test_client_2 <- predict(stades_bsr, new_data = new_tibble_team_2, type = "class")
new_tree_1 <- predict(stades_tree_1, new_data = new_tibble_team_2, type = "class")
new_tree_1_prob <- predict(stades_tree_1, new_data = new_tibble_team_2, type = "prob")

bar_prob  <- round(pred_test_client_2$.pred_yes * 100, 2)
bar_class <- as.character(class_test_client_2$.pred_class)
bar_tree_class <- as.character(new_tree_1$.pred_class)
bar_tree_prob  <- round(new_tree_1_prob$.pred_yes * 100, 2)

tibble(
  Model = c("Logistic regression", "Classification tree"),
  Prediction = c(bar_class, bar_tree_class),
  `Probability of profit` = c(paste0(bar_prob, "%"), paste0(bar_tree_prob, "%"))
) %>%
  knitr::kable(caption = "Bar scenario prediction") %>%
  plain_table()
```

Switching the café to a bar raises the estimated probability of profit to `r bar_prob`%. 

# 16. Key Findings

- The current kiosk scenario is classified as not profitable by both models.
- Its logistic-model probability of profit is `r current_prob`%.
- A vending machine lowers the modelled probability to `r vending_prob`%.
- A bar raises the modelled probability to `r bar_prob`%, with both models predicting profit.
- The profitable-class ROC specification produces an AUC of `r auc_pct`%.



# 17. Conclusion

The selected logistic model gives useful but imperfect discrimination between profitable and not-profitable club seasons. With `yes` explicitly defined as the profitable event, the held-out results show `r acc_pct`% accuracy, `r sens_pct`% sensitivity, `r spec_pct`% specificity, `r prec_pct`% precision, and an AUC of `r auc_pct`%.

The current kiosk configuration remains at a `r current_prob`% predicted probability of profit. The bar scenario produces the highest predicted probability among the tested café formats, at `r bar_prob`%. So, it will be the efficient decision for the club to change their cafe style into bar. That's the fast and furious decision to take a chance to lessen the chances of loss! But this is a conditional scenario comparison according to data what we have in our hand. A costed business case is required before implementation.


# 18. Risks and Limitations

- Observational data does not establish that café format causes profitability.
- The source clubs may differ from the French operating context.
- Profitable-class sensitivity is stronger than not-profitable specificity, so false-positive classifications remain a material risk.


# 19. Technical Appendix

## 19.1 Variable Definitions

- `profit`: `yes` when revenue exceeds costs, `no` otherwise.
- `ade`: average years of player experience.
- `members`: number of paying members.
- `coupe`: recent premiership indicator.
- `cond`: stadium condition.
- `cafe`: café service type.

## 19.2 Model Geometry

The model contains two quantitative predictors, `ade` and `members`, and categorical structures for premiership history and café type. The retained interactions allow the membership slope to vary across premiership and café categories.

## 19.3 Scenario Inputs

All three stadium scenarios hold experience at \(40 years with 7 players\), stadium condition at poor, premiership history at no, and membership at 1,500. Only café type changes between scenarios.

## 20.4 Reproducibility

This report uses a fixed random seed, an 18,000/7,000 train/test split, and portable relative file paths throughout. `tidyverse` supports data preparation. `tidymodels` supports modelling. `yardstick` provides explicitly oriented classification and ROC metrics. `rpart.plot` visualises trees. `vip` plots variable importance. `car` supplies analysis-of-deviance testing.

This project originated from university coursework and has been reorganised as an analytics portfolio case study.
