---
title: "Student Performance Analytics: Exploratory Analysis"
subtitle: "Understanding Student Characteristics, Academic Outcomes, and Initial Relationships"
author: "Md Soad Solaiman"
format: html
editor: visual
---

## Executive Summary

This report examines student performance through descriptive statistics, distribution analysis, and comparisons across tutoring groups. The dataset contains 2,392 high school student records covering demographics, study habits, parental involvement, extracurricular activities, and academic outcomes.

Within this sample, 721 students received tutoring and 1,671 did not. Students receiving tutoring exhibited higher median and mean GPAs and larger proportions of A and B grades. These comparisons describe associations in the observed sample and do not establish that tutoring caused the differences.

The GPA distribution is broadly balanced around its centre, with a mean of approximately **1.91** and a median of approximately **1.89**. A Normal model provides a useful approximation for the central portion of the distribution, although the Q-Q plot shows departures in both tails. The Normal-model estimate of the proportion below a GPA of 2.5 is approximately **74.2%**, compared with an observed proportion of approximately **70.5%**. The corresponding 90th-percentile estimates are approximately **3.08** under the Normal model and **3.13** empirically.

## Dataset Overview

The analysis uses the repository dataset at `data/student-performance-data.csv`. It contains one record per student and includes demographic characteristics, study behaviours, extracurricular participation, parental involvement, GPA, and grade classification.

The exploratory workflow focuses on tutoring participation, GPA, and grade classification. It summarises individual distributions, compares outcomes across tutoring groups, and assesses how well a Normal distribution approximates GPA for probability and percentile calculations.

## Variable Classification

| Variable        | Variable classification | R data representation |
|-----------------|-------------------------|-----------------------|
| Age             | Quantitative discrete   | Integer               |
| Gender          | Categorical nominal     | Factor                |
| StudyTimeWeekly | Quantitative continuous | Double                |
| Absences        | Quantitative discrete   | Integer               |
| GPA             | Quantitative continuous | Double                |
| GradeClass      | Categorical ordinal     | Ordered factor        |

**Age** is a discrete quantitative variable recorded in whole years from 15 to 18 and can be represented as an integer.

**Gender** is a nominal categorical variable coded as 0 for male and 1 for female. For analysis and presentation, these codes are appropriately represented as factor levels.

**StudyTimeWeekly** is a continuous quantitative variable representing weekly study time from 0 to 20 hours. Its decimal values are represented as doubles.

**Absences** is a discrete quantitative variable counting absences from 0 to 30 and can be represented as an integer.

**GPA** is a continuous quantitative measure on a scale from 0 to 4. Its decimal values are represented as doubles.

**GradeClass** is an ordinal categorical variable coded from 0 to 4, corresponding to grades A through F. The codes identify ordered categories rather than numerical magnitudes and are therefore represented as factor levels for analysis.

## Exploratory Data Analysis

### Tutoring Participation

```{r}

student_data <- read.csv("data/student-performance-data.csv")


student_data$Tutoring <- factor(student_data$Tutoring,
                                levels = c(0, 1),
                                labels = c("No Tutoring", "Tutoring"))

table(student_data$Tutoring)
```

```{r}

tutoring_counts <- table(student_data$Tutoring)

barplot(tutoring_counts,
        main = "Student Participation in Tutoring",
        xlab = "Tutoring Status",
        ylab = "Number of Students",
        ylim = c(0, 2000),
        yaxp = c(0, 2000, 4),
        col = c("#2563A6", "#0F766E"),
        border = "black")
```

Tutoring participation was less common in the sample. A total of 1,671 students did not receive tutoring, while 721 received tutoring. The non-tutoring group was therefore more than twice the size of the tutoring group.

## Distribution Analysis

### GPA Distribution

```{r}
hist(student_data$GPA,
     main = "Distribution of Student GPA",
     xlab = "Grade Point Average",
     ylab = "Number of Students",
     col = "#0F766E",
     border = "black",
     breaks = 50,
     xlim = c(0, 4),
     ylim = c(0, 120),
     yaxp = c(0, 120, 6),
     cex.main = 1.3,
     cex.lab = 1.2,
     cex.axis = 1.1)
```

```{r}

gpa_summary <- data.frame(
  Statistic = c("Mean", "Standard Deviation", "Median", "IQR"),
  Value = c(
    mean(student_data$GPA),
    sd(student_data$GPA),
    median(student_data$GPA),
    IQR(student_data$GPA)
  )
)

gpa_summary


```

The histogram indicates a broadly balanced distribution around the centre, while the similarity of the mean and median suggests limited central skew. These observations do not imply perfect symmetry or exact Normality.

```{r}

boxplot(student_data$GPA,
        main = "GPA Distribution and Potential Outliers",
        sub = "Based on 2,392 High School Students",
        ylab = "Grade Point Average",
        xlab = "All Students",
        ylim = c(0, 4.5),
        col = "#DCEAF4",
        border = "#17324D",
        outline = TRUE,
        cex.main = 1.3,
        cex.lab = 1.1,
        cex.axis = 1)

```

The sample has a mean GPA of approximately **1.91**, a standard deviation of approximately **0.92**, and a median of approximately **1.89**. The close mean and median support the use of the mean and standard deviation as useful summaries of centre and spread. The boxplot does not identify prominent observations beyond its standard plotting rule, although this visual result should not be interpreted as proof that unusual observations are absent.

## Relationships Between Variables

### Tutoring Status and Grade Classification

```{r}

student_data$GradeClass <- factor(student_data$GradeClass,
                                  levels = c(0, 1, 2, 3, 4),
                                  labels = c("A", "B", "C", "D", "F"))


tutoring_grade_table <- table(student_data$Tutoring, student_data$GradeClass)
tutoring_grade_table
```

```{r}

prop.table(tutoring_grade_table, margin = 1)
```

```{r}

tutoring_grade_pct <- prop.table(tutoring_grade_table, margin = 1) * 100

barplot(tutoring_grade_pct,
        beside = TRUE,
        main = "Grade Classification by Tutoring Status",
        xlab = "Grade Classification (A = Highest, F = Lowest)",
        ylab = "Students Within Tutoring Group (%)",
        col = c("#2563A6", "#C65D21"),
        border = "black",
        ylim = c(0, 65),
        yaxp = c(0, 60, 6),
        space = c(0.1, 0.6),
        cex.main = 1.3,
        cex.lab = 1.2,
        cex.axis = 1.1,
        cex.names = 1.1,
        legend.text = c("No Tutoring", "Tutoring"),
        args.legend = list(title = "Tutoring Status",
                           x = "topleft",
                           bty = "n",
                           cex = 1.0))


```

Grade distributions differed between the two tutoring groups. An A grade was recorded for **7.2%** of students receiving tutoring, compared with **3.3%** of students not receiving tutoring. The tutoring group also had a larger proportion of B grades.

An F grade was recorded for **45.6%** of students receiving tutoring and **52.8%** of students not receiving tutoring. These sample differences indicate an association between tutoring status and grade classification. They do not demonstrate a causal effect because the analysis does not control for other characteristics that may differ between the groups.

### GPA Across Tutoring Groups

```{r}

boxplot(GPA ~ Tutoring, 
        data = student_data,
        main = "GPA Distribution by Tutoring Status",
        xlab = "Tutoring Status",
        ylab = "Grade Point Average",
        col = c("#2563A6", "#C65D21"),
        border = "black",
        notch = FALSE,
        outpch = 19,
        outcol = "gray30")

# Summary statistics by group
tapply(student_data$GPA, student_data$Tutoring, summary)

```

Students receiving tutoring had a median GPA of approximately **2.10**, compared with approximately **1.82** among students not receiving tutoring. The corresponding mean GPAs were approximately **2.11** and **1.82**. The boxes are similar in size, indicating broadly comparable within-group dispersion.

The observed sample therefore shows higher central GPA values among students receiving tutoring. This is a descriptive comparison only; it does not account for potential confounding variables or establish that tutoring caused the difference.

## Normality Assessment

```{r}
qqnorm(student_data$GPA,
       main = "Normal Q-Q Plot of Student GPA",
       xlab = "Theoretical Normal Quantiles",
       ylab = "Observed GPA Quantiles",
       pch = 19,
       col = "#0F766E",
       cex = 0.8,
       cex.main = 1.3,
       cex.lab = 1.2,
       cex.axis = 1.1)

qqline(student_data$GPA,
       col = "#C65D21",
       lwd = 2)
```

The central observations in the Q-Q plot follow the reference line relatively closely, particularly between theoretical quantiles of approximately −1 and 1. Departures are visible in both tails, indicating that the full GPA distribution is not exactly Normal. The Normal distribution provides a reasonable approximation for the central portion of the data, but probability and percentile estimates derived from it should be compared with empirical results and interpreted with caution.

## Probability Analysis

Using the sample mean and sample standard deviation as parameter estimates, the analysis compares a Normal-model probability with the corresponding observed proportion.

```{r}

gpa_mean <- mean(student_data$GPA)
gpa_sd <- sd(student_data$GPA)


prob_gpa_less_2_5 <- pnorm(2.5, mean = gpa_mean, sd = gpa_sd)

empirical_prob <- sum(student_data$GPA < 2.5) / nrow(student_data)


probability_table <- data.frame(
  Measure = c("Sample Mean (GPA)",
              "Sample SD (GPA)",
              "P(GPA < 2.5) [Normal Model]",
              "P(GPA < 2.5) [Empirical Proportion]"),
  Value = c(gpa_mean, gpa_sd, prob_gpa_less_2_5, empirical_prob)
)

probability_table
```

Under the fitted Normal model, the probability of a GPA below 2.5 is approximately **0.742**, or **74.2%**. The observed proportion is approximately **0.705**, or **70.5%**. Both measures indicate that most students in the sample have a GPA below 2.5, but the Normal model estimates a proportion approximately 3.7 percentage points higher than the empirical calculation. This difference is consistent with the tail departures observed in the Q-Q plot.

## Percentile Analysis

```{r}


normal_cutoff <- qnorm(0.90, mean = gpa_mean, sd = gpa_sd)

empirical_cutoff <- quantile(student_data$GPA, 0.90)


cutoff_table <- data.frame(
  Method = c("Normal Model (90th percentile)", "Empirical (90th percentile)"),
  GPA_Cutoff = c(normal_cutoff, empirical_cutoff)
)

cutoff_table
```

The Normal-model estimate of the 90th-percentile GPA is approximately **3.08**. The empirical 90th percentile is approximately **3.13**, meaning that 90% of observed GPAs are at or below this value under the empirical quantile definition. The estimates are close, but the empirical cutoff is approximately 0.05 GPA points higher. The difference should be understood as a model-versus-data discrepancy rather than evidence of a specific form of skewness.

## Key Findings

- Tutoring participation was less common than non-participation, with 721 students receiving tutoring and 1,671 not receiving tutoring.
- The sample mean GPA was approximately **1.91**, the median was approximately **1.89**, and the standard deviation was approximately **0.92**.
- Students receiving tutoring exhibited higher mean and median GPAs and larger proportions of A and B grades within the sample.
- The central portion of the GPA distribution was reasonably approximated by a Normal distribution, while both tails departed from the Normal reference.
- The Normal model estimated that **74.2%** of students had a GPA below 2.5, compared with an observed proportion of **70.5%**.
- The Normal-model and empirical 90th-percentile GPAs were approximately **3.08** and **3.13**, respectively.

## Limitations

The analysis is descriptive and does not establish causal relationships. Comparisons by tutoring status do not adjust for study time, attendance, parental support, prior performance, or other potential confounders. The sample design is not evaluated here, so the extent to which findings generalise beyond these records is uncertain.

GPA is bounded between 0 and 4, whereas a Normal distribution is unbounded. Although the Normal model provides a useful central approximation, the visible tail departures and differences from empirical estimates show that it does not fully represent the observed distribution. Probability and percentile estimates should therefore be interpreted alongside their empirical counterparts.
